---
title: "Accidents Crash Analysis"
subtitle: "Proposal"
author: 
  - name: " Team Data Dazzlers"
    affiliations:
      - name: "School of Information, University of Arizona"
description: "Accidents Crash Analysis"
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
```

## Dataset

```{r}
#| label: load-dataset
#| message: false
library(here)
crash <- read.csv(here('data','NYC_Motor_Vehicle_Collisions_to_Person.csv'))
crash|>glimpse()
crash
```

## Description of the Dataset:

The dataset "NYC_Motor_Vehicle_Collisions_to_Person," sourced from the official website of the Massachusetts Registry of Motor Vehicles, will be utilized for this project. This dataset was selected for its relevance in the context of a worrying increase in road accidents over the past 30 years. Analyzing this data could reveal patterns that would enable the creation of algorithms or models capable of predicting crashes. Such predictive tools could significantly aid in preventing accidents and potential fatalities. By examining this data, we aim to contribute by identifying the places and times where accidents are likely to occur, thereby enhancing public safety and ensuring safer travel experiences.

The dataset comprises 72 dimensions, including Crash Number, City/Town Name, Crash Date, Crash Severity, Crash Time, Crash Year, Maximum Injury Severity Reported, Age of the Youngest Known Driver, and Age of the Oldest Known Driver, among others.

## Questions

The two questions you want to answer.

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).

**Question 1: What age group has been involved in the accidents and at what time the accidents were at peak?** 

**Approach**

**Plots:**

1\) plot 1:

2\) plot 2:

\
The second plot can rely on a time series or line plot for accidents over time of day. Therefore, through aggregating the number of accidents by hour, the plot can visualize the trend of accidents throughout a day. It is also suitable for using Shiny app which allow users to interact with the plot. There can be a time series bar, when the user drag the button, the plot can show different numbers of accidents in different hours in a day. Also, this plot can help identify peak times for accidents. \

**Variables to be used and created:**

The main variables will be used to answer this question are: 

1.  **Age**: which will help to categorize individuals into different groups to see which age group is more involved in accidents

2.  **Crash Time:** which will help to determine the times when accidents are most frequent. 

The variables will be created have not yet be determined, but some variables can be considered are: 

1.  **Age Group:** which can categorize individuals into specific age groups, such as 0-18 (youths), 19-35 (young adults), 36- 50 (middle-aged adults), 51-65 (pre-seniors) and 65+ (seniors). There are columns with age groups, but the age can be regrouped so they can be grouped with different categories which allow for a clearer analysis of which age groups seem to be involved in accidents

2.  **Day Part**: instead of looking at the specific time,  it is helpful to divide day into different parts like Morning (5am-12:00am), Afternoon (12:01pm - 5:00pm), Evening (5:01pm - 9:00 pm), and Night (9:01 pm - 4:59am). In such a way, not only peak hours but also period of time can used to see when more accidents happen. 

\
\
